\huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation

### 2019-07-26 Manuscript_Publication.Rmd
More writing

---
title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
  in birds}
author:
- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
  and Iliana Medina Guzmán\textsuperscript{§}
- '*\textit{justin.g.cally@gmail.com}'
bibliography: References/SS_Extinction.bib
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/evolution.csl
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates. However, many of the hypothesised effects of sexual selection are contradictory, and the net effect of sexual selection on diversification is unclear. Here, we investigate the relationship between sexual selection and diversification in passerine birds, using two measures of sexual selection --- sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny and male parental care. We also test whether the relationship between sexual selection on diversification is contingent on ecological variables, as some theories predict. Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines. We also find that birds with smaller ranges have higher speciation rates, but there is no interaction between environmental variables and sexual selection. Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds. \vspace{5mm}  \par\noindent
  \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

@@

Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008]. There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and diversification [e.g. @Lande_1981; @Lande_1982]. Experimental work suggests that sexual selection shapes the evolution of many different traits --- not only the obvious ones such as courthship songs or antlers [@Cally_2019] --- while some evolutionary comparative analyses have found that variation in sexual selection between clades predicts phylogenetic diversity. For instance, in antbirds (Thamnophilidae), genera with complex songs and pronounced sexual dichromatism are more species-rich [@Seddon_2008]. ADD ONE MORE EXAMPLE IF POSSIBLE

Several hypotheses predict a positive relationship across clades between the strength of sexual selection and the diversification rate (or the amount of extant phylogenetic diversity). For example, sexual selection is the mechanism underlying reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REFERENCES**). Additionally, sexual selection often operates on traits that can create reproductive isolation when they differ between lineages, such as sexual signals and preferences, sperm-egg interactions, or genitalia (**Safran_2013?, MORE - @Lande_1981; @Lande_1982 ?**). Furthermore, sexual selection is hypothesised to increase diversification by maintaining trait combinations that would be disadvantageous under pure natural selection (**REFS - Bonduriansky_2011 & Radwan_2016? NB I think the journal name might be wrong for Radwan_2016, it's probably JEB**), and by helping to prevent extinction by purging deleterious mutations [@Agrawal_2001; @Siller_2001], fixing beneficial ones [@Whitlock_2000], and accelerating the rate at which populations can adapt to novel environments [@Lorch_2003; @Candolin_2008; @Cally_2019]. Also, sexual selection is sometimes diversifying (as opposed to directional or stabilising), for example when individuals with rare phenotypes have elevated mating success (**REFS? - Holman et al BMC Evol Biol 2009 about Crozier's paradox is one, and likely has some refs in it**): diversifying selection promotes speciaiton by increasing the rate at which isolated populations become genetically distinct.

Conversely, there are multiple ways in which sexual selection is hypothesised to reduce the diversification rate. Firstly, sexual selection can make extinction more likely by favouring traits that improve mating success but reduce population fitness. For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the mating success of individuals while reducing population-wide productivity [e.g. infanticide, harassment, and reduced parental care in mate-seeking males; @RN37; @Holman_Kokko_2013]. Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013] **Add paper with Mikatake, Hosken and other. Current Biology, about selection on a male trait in beetles constraining female trait evolution even tho the male trait is male-only**. Secondly, sexual selection might decrease the speciation rate by ensuring that individuals with novel trait values have low reproductive success.

Importantly, the relationship between sexual selection and diversification may depend on the environment. Some theoretical and empirical results suggest that sexual selection has a more positive effect on adaptation and population fitness in variable relative to stable environments [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017; @Cally_2019]. The hypothesised mechanism is that stable environments lead to the erosion of genetic variation at sexually concordant loci (*i.e.* loci where the fittest allele is the same in both sexes), while sexually antagonistic loci remain comparatively genetically variable, reducing the genetic correlation between male and female fitness. In turn, this means that the most attractive members of the more competitive sex (typically males) produce less fit offspring of the demographically important sex (usually female), such that sexual selection is less likely to favour female-detrimental, extinction-promoting alleles in variable environments. Furthermore, sexual selection might be especially helpful in preventing extinction in variable environments, because (in the words of Charles Darwin), "sexual selection will give its aid to ordinary selection, by assuring to the most vigorous and best adapted males the greatest number of offspring" (DARWIN REF; see also @Agrawal_2001; @Siller_2001; @Boughman_2002; @Maan_2011; ALSO KOKKO + HEUBEL 2008 model about GxE).

@@

Birds have been a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003]. Birds are well-suited for this questions as their phylogenetic relationships are comparateively well-known, and they have diverse mating systems and conspicuous sexually-selected traits such as song and colourful plumage. A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011]. However, the meta-analysis found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.

Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways. Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual size dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001]. In fact, it was recently highlighted that song divergence is inversely correlated with sexual dichromatism in passerines, suggesting that a focus on one signalling modality may cause the mismeasurement sexual selection [@Cooney_2018]. Our study compares the role of sexual dichromatism and other sexual selection proxies in the **generation of species (Really? Isn't the composite of speciation and extinction?)** . Additionally, we test the relative effect of some ecological and environmental variables **(I say "ecological" because of range size)** on diversification, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].

To investigate the link between different measures of sexual selection and diversification, we draw upon a large (*n* = 5,812 species) dataset of sexual dichromatism [estimated from illustrations; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that provides an index of male-biased sexual selection, which uses principal components analysis to capture (co)variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015]. These datasets allow us to cover the majority of bird genera. Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$). We also check the robustness of the results to phylogenetic uncertainty, by considering the posterior distribution of possible phylogenies rather than restricting our analysis to the single most likely phylogeny **(REF?)**.  \phantomsection

@@

**Is it possible to move this later in the Results? It's maybe a bad idea to start with confusing, worrying results that don't pertain to our main non-null finding regarding the SS index**

We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees from the posterior distribution of trees [@Jetz_2012]. Estimates of speciation rates varied between phylogenetic trees, especially for rates calcualted by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the 95% confidence interval calculated for the MCC tree. Conversely, for other **DIVERSIFICATION??** rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar likelihoods is notable. Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Details of the BAMM results are presented in the supplementary materials.

\phantomsection
## Effects of sexual dichromatism/selection on speciation do not depend on environmental variability

**I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Intro.**

We next fit PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection. However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

@@

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$. The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

@@

![Path analysis of evolutionary and ecological variables. Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative). The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero. The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465). \label{Path_Analysis}](Figures/Path_Analysis.pdf)

>>>>

---
title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
  in birds}
author:
- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
  and Iliana Medina Guzmán\textsuperscript{§}
- '*\textit{justin.g.cally@gmail.com}'
bibliography: References/SS_Extinction.bib
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/evolution.csl
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates. However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes. Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera). We tested for associations between two measures of sexual selection (sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny, and male parental care), several environmental variables, and multiple measures of speciation and extinction rates. Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies with the environment. We also found a strong negative relationship between range size and speciation rate. There was no correlation between sexual selection and extinction rate **XXXXXX**. Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds. \vspace{5mm}  \par\noindent
  \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

<!-- Luke general comments:
In a few places you say things like "1000 random trees". To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al. Is that right? If so, probably best to search for the word "random" and remove/reword it. 
-->

@@

Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and in most species it predominently operates on males [@Kokko_Jennions2008; @Fromhage_2016; @Janicke_2016]. There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and phylogenetic diversification [e.g. @Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018]. Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g. life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods (**REFERENCES**), present new opportunities to test whether and how sexual selection drives diversification.

Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction. For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REF**). Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia]. Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019]. Also, sexual selection sometimes favours individuals with rare phenotypes [@holman2013crozier], providing a source of diversifying selection that might increase the rate at which isolated populations diverge genetically.

Conversely, there are multiple ways in which sexual selection might hinder speciation or make extinction more likely. Firstly, sexual selection might promote extinction by favouring traits that improve mating success but reduce population fitness. For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g. infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016]. Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013]. Secondly, sexual selection might promote evolutionary stasis by ensuring that individuals with novel trait values have low mating success.

Importantly, the relationship between sexual selection and diversification might be contingent on the environment. Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [e.g. @Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019]. The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (*i.e.* loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex). In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g. @Connallon_Hall_2016], and potentially to a more negative relationship between sexual selection and diversification rates across species. Conversely, sexual selection might be especially important at promoting local adaptation when the environment displays variability at the relevant spatial and temporal scales [e.g. @Boughman_2002; @Lorch_2003; @Maan_2011; @kokko2008condition]

@@

Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [*e.g.* @Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014]. Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known. A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011]. However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.

Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways. We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care. Most **OR ALL??** earlier macroevolutionary studies of this topic only quantified sexual selection using dichromatism, which is problematic since dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015]. For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re]. Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018]. Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016]. We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$). Finally, we incorporate phylogenetic uncertainty into our analyses, improving over earlier work which assumed the statistically best-fitting phylogeny is correct.  \phantomsection

@@

**[Is it possible to move this later in the Results? It's possibly not ideal to start by highlighting the dependence of our conclusions on the methods we choose, and not mentioning our main non-null finding regarding the SS index. Having said that I can see a case for leaving it here too. - Luke]**

We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees sampled from the posterior distribution of trees in @Jetz_2012. Estimates of speciation rates varied between phylogenetic trees, especially for rates calculated by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the **95% confidence interval [SHOULD THIS BE HPD TOO?]** calculated for the MCC tree. Conversely, for our estimates of diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95% confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable. Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Full details of the BAMM results are presented as supplementary materials.

\phantomsection
## No evidence that the relationship between sexual selection and speciation depends on environmental variability

**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro. - Luke]**

We next fit Phylogenetic Generalized Least Squares (PGLS) models with **2-way [CORRECT?]** interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection. However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

@@

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$. The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

@@

![Path analysis of evolutionary and ecological variables. Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative). The numeric values are standardised regression slopes and the asterisks indicates that the 95% confidence intervals of this estimate do not overlap with zero. The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465). \label{Path_Analysis}](Figures/Path_Analysis.pdf)
